---
title: "gtsummary in quarto"
format: html
execute: 
  echo: false
theme: quartz
---
The NLSY79 Cohort is a longitudinal project that follows the lives of a sample of American youth born between 1957-1964. The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979; after two subsamples were dropped, 9,964 respondents remain in the eligible samples. Data are now available from Round 1 (1979 survey year) to Round 30 (2022 survey year).

```{r}
#| message: false

library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.txt"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

Exercise Part 1

```{r}
#| label: tbl-sex
#| tbl-cap: "Descriptive Statistics"

tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

Descriptive statisitcs from the NLSY79 Cohort (@tbl-sex)

```{r}
mean_sleep_wknd <- mean(nlsy$sleep_wknd, na.rm = TRUE)
```
Participates slept an average of `r mean_sleep_wknd` hours on the weekend.

```{r}
median_sleep_wknd <-median(nlsy$sleep_wknd, na.rm = TRUE)
```

The median hours of sleep on weekends was `r median_sleep_wknd`.

Exercise Part 2

```{r}
#| label: tbl-descr
#| tbl-cap: "Descriptive statistics"
#| output-location: slide
table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir)) |> 
  add_overall(last = TRUE)
table1
```

```{r}
female_age_bir <- inline_text(table1, variable = "age_bir", column = "stat_2")
```

The median (IQR) age of the female participants at the birth of their first child was `r female_age_bir`.

```{r}
#| label: tbl-sex-sleep
#| tbl-cap: "Descriptive statistics by sex with sleep"

table_sex_sleep <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sleep_wknd),
  label = list(
  	sleep_wknd ~ "Sleep on Weekend")
  )|> 
  add_overall(last = TRUE)

table_sex_sleep

overall_sex_sleep <- inline_text(table_sex_sleep, variable = "sleep_wknd", column = "stat_0")
```
The median (IQR) hours of sleep on weekend in all participants was `r overall_sex_sleep`.

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```


